{\bf Landslide: A New Race-Finding Tool for 15-410}

### 2016-09-28 landslide-lecture/2016/lecture.tex
import 2015 landslide lecture to new 2016 dir
Update lecture for f16

\institute[CMU 15-410]{Carnegie Mellon University - 15-410}
\date[]{2016, February 10}

@@

\begin{frame}{Testing Mechanisms}
	\textbf{Stress testing}: \texttt{largetest}, \texttt{mandelbrot} and friends
	\begin{itemize}
		\item Attempting to exercise as many interleavings as practical
		\item Exposes race conditions at random
		\begin{itemize}
			\item ``If a preemption occurs at just the right time\ldots''
		\end{itemize}
		\item Cryptic panic messages when failure occurs
	\end{itemize}
	\linegap
	What if\ldots
	\begin{itemize}
		\item Make educated guesses about when to preempt
		\item Preempt enough times to run {\em every single} interleaving
		\item Tell the story of what {\em actually happened}.
		\item Overlook fewer bugs!
	\end{itemize}
	% Say: At the end of this lecture, I'm going to be looking for some among you...
\end{frame}

@@

\begin{frame}{Systematic Testing - The Big Picture}
	\textbf{Goal: Force the system to execute every possible interleaving.}
	% Say: checking for crashes/memory errors/etc in each one.
	\begin{itemize}
		\item On 1st execution, schedule threads arbitrarily until program ends.
			\begin{itemize}
				\item This represents one branch of the tree.
			\end{itemize}
		\item At end of each branch, rewind system and restart test.
		\item Artificially add preemptions to produce different thread interleavings.
		\item Intuitively: Generate many ``tabular execution traces''.
	\end{itemize}
	\pause
	\linegap

@@

	Systematic tests are {\em parameterized} by the set of PPs.
	\begin{itemize}
		\item If there are $n$ PPs and $k$ threads, state space size is $n^k$.
		\item Need to choose the set of PPs very carefully for test to be effective.
		\begin{itemize}
			\item ``Effective'' = both comprehensive and feasible.
		\end{itemize}
	\end{itemize}
\end{frame}

@@

	\linegap
	When do we fear being preempted?
	\begin{itemize}
		\item Threads becoming runnable (\texttt{thr\_create()}, \texttt{cond\_signal()}, etc.)
			\begin{itemize}
				\item Preemptions may cause it to run before we're ready
			\end{itemize}
		\item Synchronization primitives (\texttt{mutex\_lock()}/\texttt{unlock()}, etc.)
			\begin{itemize}
				\item If buggy or used improperly\ldots
			\end{itemize}
		\item Unprotected shared memory accesses (``data races'')% Say: more on this later.
			\begin{itemize}
				\item May result in data structure corruption
				\item More on this later...
			\end{itemize}
	\end{itemize}

@@

\begin{frame}{About The Project}
	\textbf{About me}: 5th year graduate student, advised by Garth Gibson
	\begin{itemize}
		\item TAed 15-410 for 3 semesters during undergrad
		\item 1st graduate year was 5th year M.S., rest in Ph.D. program
		\begin{itemize}
			\item \url{http://www.contrib.andrew.cmu.edu/~bblum/thesis.pdf}
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

	{\bf About Landslide}
	\begin{itemize}
		\item Simics module, which traces:
			\begin{itemize}
				\item Every instruction executed
				\item Every memory access read/written
			\end{itemize}
			% FIXME for next year -- how much slowdown is there? 20x?
		\item Originally supported only P3s; can now test P2s fully-automated
		\item {\em Landslide} shows how your {\em Pebbles} programs may not be stable.
	\end{itemize}
\end{frame}

@@

	State of the art tools hard-code a fixed set of preemption points.
	\begin{itemize}
		\item E.g., ``all thread library API calls'' or ``all kernel mutex locks/unlocks''
		\item Depending on length of test, completion time is unpredictable.
		\item More often, a subset is better in terms of time/coverage. 
	\end{itemize}
	\pause
	\linegap
	Current systematic testing model is not user-friendly.
	\begin{itemize}
		\item Tool: ``I want to use these PPs, but can't predict completion time.''
		\item User: ``I have 16 CPUs and 24 hours to test my program.''
	\end{itemize}
	\linegap
	{\em Stress testing allows user to choose total run time -- can we offer this too?}
\end{frame}

\begin{frame}{Iterative Deepening of Preemption Points}
	Goal: Run the best tests for a given CPU budget.
	\begin{itemize}
		\item Technique: ``Iterative Deepening''
	\end{itemize}
	\linegap

@@

\begin{frame}{Last Semester}
	%Spring 2015: 7 groups signed up to use Landslide; 5 found bugs
	%%\begin{itemize}
	%%	\item (that they didn't find during stress testing)
	%%\end{itemize}
	%%%%% hopefully!
	%\linegap

	%Among all groups, 38 total tests were run
	%\begin{itemize}
	%	\item 21 tests ran to completion without finding bugs
	%	\item 4 {\em deterministic} bugs
	%	\item 13 distinct {\em non-deterministic} bugs
	%		\begin{itemize}
	%			\item 6 in {\tt thr\_exit\_join} \small (use-after-free, deadlock, NULL, infinite loop...)
	%			\item 4 in {\tt paraguay} (deadlock, infinite loop)
	%			\item 2 in {\tt broadcast\_test} (deadlock, infinite loop)
	%			\item 1 in {\tt rwlock\_downgrade\_read\_test} (use-after-free)
	%		\end{itemize}
	%		%\begin{itemize}
	%		%	\item 6 of which required data-race PPs to expose
	%		%\end{itemize}
	%\end{itemize}
	Fall 2015: 22 groups signed up to use Landslide; 21 found bugs
	% 18 groups found races
	\begin{itemize}
		\item (that they didn't find during stress testing)
	\end{itemize}
	%%%% hopefully!
	\linegap

	Among all groups, 185 distinct tests were run
	\begin{itemize}
		\item 137 tests ran to completion without finding bugs
		\item 21 {\em deterministic} bugs
	\end{itemize}
	\pause % ask: "who can tell me why landslide might find a *deterministic* bug not found during past testing? A: use-after-free
	\linegap

	38 distinct {\em non-deterministic} bugs
	\begin{itemize}
			%\begin{itemize}
			%	\item 7 of which required data-race PPs to expose
			%\end{itemize}
		\item 11 groups (50\%) fixed all their bugs
			\begin{itemize}
				\item (as verified by running Landslide again -- not a guarantee!)
			\end{itemize}
		\item 2.7 bugs ``fixed'' (see above) on average per group
		\item Most ambitious group: 11 distinct races found + fixed!
	\end{itemize}
\end{frame}

@@

	To participate\dots
	\begin{itemize}
		\item Review this lecture and study information sheet
		\item Meet prerequisites of passing P2 tests
		\item Complete sign-up form online to get further instructions
			\begin{itemize}
				\item \url{http://tinyurl.com/landslide-p2-s16}
			\end{itemize}
		\item Optional ``Landslide clinic'' for in-person tech (or moral) support
			\begin{itemize}
				\item Mid-next-week, room and time TBD
			\end{itemize}
		\item Use Landslide in addition to stress tests until P2 is due!
	\end{itemize}
	\linegap

@@

\begin{frame}{Last-last Semester}
	Spring 2015: 7 groups signed up to use Landslide; 5 found bugs
	%\begin{itemize}
	%	\item (that they didn't find during stress testing)
	%\end{itemize}
	%%%% hopefully!
	\linegap

	Among all groups, 38 total tests were run
	\begin{itemize}
		\item 21 tests ran to completion without finding bugs
		\item 4 {\em deterministic} bugs
		\item 13 distinct {\em non-deterministic} bugs
			\begin{itemize}
				\item 6 in {\tt thr\_exit\_join} \small (use-after-free, deadlock, NULL, infinite loop...)
				\item 4 in {\tt paraguay} (deadlock, infinite loop)
				\item 2 in {\tt broadcast\_test} (deadlock, infinite loop)
				\item 1 in {\tt rwlock\_downgrade\_read\_test} (use-after-free)
			\end{itemize}
			%\begin{itemize}
			%	\item 6 of which required data-race PPs to expose
			%\end{itemize}
	\end{itemize}
\end{frame}

>>>>

\institute[CMU 15-410]{Carnegie Mellon University - 15-410}
\date[]{2016, September 28}

@@

\begin{frame}{Testing Mechanisms}
	\textbf{Stress testing}: \texttt{largetest}, \texttt{mandelbrot} and friends
	\begin{itemize}
		\item Attempting to exercise as many interleavings as practical
		\item Exposes race conditions at random
		\begin{itemize}
			\item ``If a preemption occurs at just the right time\ldots''
		\end{itemize}
		\item Cryptic panic messages when failure occurs
	\end{itemize}
	\linegap
	What if\ldots
	\begin{itemize}
		\item Make educated guesses about when to preempt
		\item Preempt enough times to run {\em every single} interleaving
		%\item Tell the story of what {\em actually happened}.
		\item Overlook fewer bugs!
	\end{itemize}
	% Say: At the end of this lecture, I'm going to be looking for some among you...
\end{frame}

@@

\begin{frame}{Systematic Testing - The Big Picture}
	\textbf{Goal: Force the system to execute every possible interleaving.}
	% Say: checking for crashes/memory errors/etc in each one.
	\begin{itemize}
		\item On 1st execution, schedule threads arbitrarily until program ends.
			\begin{itemize}
				\item This represents one branch of the tree.
			\end{itemize}
		\item At end of each branch, rewind system and restart test.
		\item Artificially preempt to interleave threads differently.
		\item Intuitively: Generate many ``tabular execution traces''.
	\end{itemize}
	\pause
	\linegap

@@

	Systematic tests are {\em parameterized} by the set of PPs.
	\begin{itemize}
		\item $n$ PPs and $k$ threads $\Rightarrow$ state space size is $n^k$.
		\item Need to choose PPs very carefully for test to be effective.
		\begin{itemize}
			\item ``Effective'' = both comprehensive and feasible.
		\end{itemize}
	\end{itemize}
\end{frame}

@@

	\linegap
	When are preemptions dangerous?
	\begin{itemize}
		\item Threads becoming runnable (\texttt{thr\_create()}, \texttt{cond\_signal()}, etc.)
			\begin{itemize}
				\item Preemptions may cause it to run before we're ready
			\end{itemize}
		\item Synchronization primitives (\texttt{mutex\_lock()}/\texttt{unlock()}, etc.)
			\begin{itemize}
				\item If buggy or used improperly\ldots
			\end{itemize}
		\item Unprotected shared memory accesses (``data races'')% Say: more on this later.
			\begin{itemize}
				\item May result in data structure corruption
				\item More on this later...
			\end{itemize}
	\end{itemize}

@@

\begin{frame}{About The Project}
	\textbf{About me}: 6th year graduate student, advised by Garth Gibson
	\begin{itemize}
		\item TAed 15-410 for 3 semesters during undergrad
		%\item 1st graduate year was 5th year M.S., rest in Ph.D. program
		\item Landslide's publication history
		\begin{itemize}
			\item Master's thesis
			\begin{itemize}
				\item \url{http://www.contrib.andrew.cmu.edu/~bblum/thesis.pdf}
			\end{itemize}
			\item Conference paper (to appear in OOPSLA this november)
			\begin{itemize}
				\item \url{http://www.contrib.andrew.cmu.edu/~bblum/oopsla.pdf}
			\end{itemize}
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

	{\bf About Landslide}
	\begin{itemize}
		\item Simics module, which traces:
			\begin{itemize}
				\item Every instruction executed
				\item Every memory access read/written
			\end{itemize}
			% FIXME for next year -- how much slowdown is there? 20x?
		%\item Originally supported only P3s; can now test P2s fully-automated
		\item {\em Landslide} shows how your {\em Pebbles} programs may not be stable.
	\end{itemize}
\end{frame}

@@

	State of the art tools choose a fixed set of preemption points.
	\begin{itemize}
		\item E.g., ``all thread API calls'' or ``all kernel mutex locks/unlocks''
		\item Depending on length of test, completion time is unpredictable.
		\item More often, a subset is better in terms of time/coverage. 
	\end{itemize}
	\pause
	\linegap
	Current systematic testing model is not user-friendly.
	\begin{itemize}
		\item Tool: ``I want to use these PPs, but can't predict completion time.''
		\item User: ``I have 16 CPUs and 24 hours to test my program.''
	\end{itemize}
	\linegap
	{\em Stress testing allows user to choose total run time -- can we offer this too?}
\end{frame}

\begin{frame}{Iterative Deepening of Preemption Points}
	Goal: Run the best tests for a given CPU budget.
	%\begin{itemize}
	%	\item Technique: ``Iterative Deepening''
	%\end{itemize}
	%\linegap

@@

\begin{frame}{Previous Semesters}
	%Spring 2015: 7 groups signed up to use Landslide; 5 found bugs
	%%\begin{itemize}
	%%	\item (that they didn't find during stress testing)
	%%\end{itemize}
	%%%%% hopefully!
	%\linegap

	%Among all groups, 38 total tests were run
	%\begin{itemize}
	%	\item 21 tests ran to completion without finding bugs
	%	\item 4 {\em deterministic} bugs
	%	\item 13 distinct {\em non-deterministic} bugs
	%		\begin{itemize}
	%			\item 6 in {\tt thr\_exit\_join} \small (use-after-free, deadlock, NULL, infinite loop...)
	%			\item 4 in {\tt paraguay} (deadlock, infinite loop)
	%			\item 2 in {\tt broadcast\_test} (deadlock, infinite loop)
	%			\item 1 in {\tt rwlock\_downgrade\_read\_test} (use-after-free)
	%		\end{itemize}
	%		%\begin{itemize}
	%		%	\item 6 of which required data-race PPs to expose
	%		%\end{itemize}
	%\end{itemize}
	S'15, F'15, S'16: 48 groups signed up to use Landslide; 38 found bugs
	% 18 groups found races
	%\begin{itemize}
	%	\item (that they didn't find during stress testing)
	%\end{itemize}
	%%%% hopefully!
	\linegap

	44 {\em deterministic} bugs (e.g. swexn, initialization)
	\linegap

	85 distinct {\em non-deterministic} bugs (among 32 groups)
	\begin{itemize}
		\item 24 groups (75\%) fixed $\ge 1$ such bug
			\begin{itemize}
				\item (as verified by running Landslide again -- not a guarantee!)
			\end{itemize}
		\item 15 groups (47\%) fixed {\em all} such bugs
		\item Most ambitious group: 11 distinct races found + fixed!
	\end{itemize}
	\linegap
	% 44 dets 85 races

	%Among all groups, 185 distinct tests were run
	%\begin{itemize}
	%	\item 137 tests ran to completion without finding bugs
	%	\item 21 {\em deterministic} bugs
	%\end{itemize}
	%\pause % ask: "who can tell me why landslide might find a *deterministic* bug not found during past testing? A: use-after-free
	%\linegap

	%38 distinct {\em non-deterministic} bugs
	%\begin{itemize}
	%		%\begin{itemize}
	%		%	\item 7 of which required data-race PPs to expose
	%		%\end{itemize}
	%	\item 11 groups (50\%) fixed all their bugs
	%		\begin{itemize}
	%			\item (as verified by running Landslide again -- not a guarantee!)
	%		\end{itemize}
	%	\item 2.7 bugs ``fixed'' (see above) on average per group
	%	\item Most ambitious group: 11 distinct races found + fixed!
	%\end{itemize}
\end{frame}

@@

	To participate\dots
	\begin{itemize}
		\item Review this lecture and study information sheet
		\item Meet prerequisites of passing P2 tests
		\item Complete sign-up form online to get further instructions
			\begin{itemize}
				\item \url{http://tinyurl.com/landslide-p2-f16}
			\end{itemize}
		\item Optional ``Landslide clinic'' for in-person tech (or moral) support
			\begin{itemize}
				\item Mid-next-week, room and time TBD
			\end{itemize}
		\item Use Landslide in addition to stress tests until P2 is due!
	\end{itemize}
	\linegap

@@

%\begin{frame}{Last-last Semester}
%	Spring 2015: 7 groups signed up to use Landslide; 5 found bugs
%	%\begin{itemize}
%	%	\item (that they didn't find during stress testing)
%	%\end{itemize}
%	%%%% hopefully!
%	\linegap
%
%	Among all groups, 38 total tests were run
%	\begin{itemize}
%		\item 21 tests ran to completion without finding bugs
%		\item 4 {\em deterministic} bugs
%		\item 13 distinct {\em non-deterministic} bugs
%			\begin{itemize}
%				\item 6 in {\tt thr\_exit\_join} \small (use-after-free, deadlock, NULL, infinite loop...)
%				\item 4 in {\tt paraguay} (deadlock, infinite loop)
%				\item 2 in {\tt broadcast\_test} (deadlock, infinite loop)
%				\item 1 in {\tt rwlock\_downgrade\_read\_test} (use-after-free)
%			\end{itemize}
%			%\begin{itemize}
%			%	\item 6 of which required data-race PPs to expose
%			%\end{itemize}
%	\end{itemize}
%\end{frame}
